### 2015-07-12 paper1/theory.tex

\section{Theory}

We start by introducing in a succinct manner the gravitational lensing
theory that is relevant for microlensing in general and for the scope
of the present paper in particular.  More detailed presentation of the
theory can be found in several references, such as
\cite{2001stgl.book.....P}.

\subsection{Microlensing essentials}

The gravitational lens equation
\begin{equation}
\vec\beta = \vec\theta - \vec\alpha(\vec\theta)
\label{eqn:lens}
\end{equation}
relates the apparent sky position $\vec\theta$ of a light source to
its true but unobservable sky position $\vec\beta$ through the bending
angle $\vec\alpha$.  The latter is an integral 
\begin{equation}
\vec\alpha = (1+z_{_L})\frac{D_S-D_L}{D_SD_L} \frac{4G}{c^2}
\int \Sigma(\vec\theta')
\frac{\vec\theta-\vec\theta'}{|\vec\theta-\vec\theta'|^2}\,
d^2\vec\theta'
\end{equation}
depending on the projected density $\Sigma(\vec\theta)$ (kg/steradian)
of the lens, the lens redshift $z_{_L}$ and the comoving distances
$D_L$ and $D_S$ to the lens and source.  The derivative of the
apparent position with respect to the source position
\begin{equation}
M(\vec\theta) =
\left(\frac{\partial\beta}{\partial\theta}\right)^{-1}
\label{eqn:magnif-matrix}
\end{equation}
is known as the magnification matrix, and its determinant
\begin{equation}
\mu(\vec\theta) = \det|M(\vec\theta)|
\end{equation}
is the brightness amplification of an image of a point source.  In
other words, the source will brighten or dim according to whether
$\mu(\vec\theta)$ is more or less than unity.  If there are multiple
mages at distinct but not observationally resolved $\vec\theta_i$ from
the same $\vec\beta$, a total brightness amplification
of\begin{equation} \mu_{\rm total} = \sum_{i} \mu(\vec\theta_i)
\end{equation}
applies. It is possible for the magnification to become formally
infinite, as a result of an eigenvalues of the magnification matrix
(\ref{eqn:magnif-matrix}) becoming infinite.  This generically happens
on curves on the $\vec\theta$ plane, known as {\em critical curves}.
Mapping a critical curve to the source plane, through the lens
equation (\ref{eqn:lens}), to the source plane $\vec\beta$ gives the
so-called caustics.  Caustics can appear in optical system, not just
gravitational lensing.  For a point source, caustics are singularities
of the magnification; for finite-size sources caustics correspond to
high and sharply changing magnification.

Caustics are important in all forms of lensing with multiple images,
but they have a special significance for lens quasars, first pointed
out by \cite{1979Natur.282..561C}.  The granularity of the mass
distribution $\Sigma(\vec\theta)$ due to individual stars produces a
caustic network on the scale of $\sqrt{GM_\odot D_L/c^2}$ or
micro-arcsec.  Extended sources wash out this micro-caustic structure,
but the optical-continuum source of quasars is even smaller.  Hence
sharp changes in the brightness of lensed quasars are observable, as
they cross the micro-caustics \citep[e.g.,][]{2012A&A...544A..62S}.

\begin{figure}
\centering
  \includegraphics[width=.9\hsize]{plots/IRIS567_path_2.eps}
\caption{\label{fig:magnification_map} Magnification map}
\end{figure}

\subsection{Magnification near a fold caustic}

A simple example of a caustic network is shown in
Figure~\ref{fig:magnification_map}.  There are two general categories
of caustics in gravitational lensing, cusps and fold, and examples of
both kinds can be seen in this figure.  Magnification near a caustic
has universal properties, independent of the system and has been
extensively studied
\citep{1986ApJ...310..568B,1992A&A...260....1S,2002ApJ...574..970G,2002ApJ...580..468G}.
In particular, at distance $p$ from a fold caustic
\begin {equation}
 \mu(p) = \mu_0 + C_0 \frac{1}{\sqrt{p}} \Theta(p).
\end {equation}
Here the magnification of a point source near a caustic is equal to
the sum of the magnification due to other reasons $\mu_0$, assumed to
be locally constant, and a decrease with the square root of the
distance from the fold. The later term becomes activated only after
the source enters the region interior to the caustic curves when the
values of the stepfunction $\Theta(p)$ become unity. The
proportionality constant $C_0$ depends on the local conditions in the
vicinity of the caustic.

A source of arbitrary shape can be described by a two dimensional brightness function $S_{2D}(p - p_s, q - q_s)$ defined for a coordinate system $p,q$ where $p_s, q_s$ denote the coordinates of the center of source.

\begin{figure}
\includegraphics[width = .49\textwidth]{plots/infinite_fold.eps}
\caption{\label{fig:infinite_fold} Source profiles for Gaussian, disk,
  crescents (right) and magnification map for an infinite fold
  (left). Objects have the same $S_{0}$ and $r_{1/2}$. }
\end{figure}

For a microlensing event the lightcurve can be written for an
undefined source shape as:
\begin{equation}
 F(t) = \int_{-\infty}^\infty \int_{-\infty}^\infty S_{2D}(p-p_s(t), q-q_s(t)) \mu_t(p) \mathrm{d}q \mathrm{d}p
 \label{eqn:ft2d}
\end{equation}
In order to build the previous equation we have considered that the
time dependency of the flux $F$ is given only by the motion of the
source with respect to a fixed caustic. Therefore the only time
dependent quantities in the right hand side of the equation are the
coordinates of the center of the source $p_s,q_s$ and by construct
$S_{2D}$.

Due to the choice of the coordinate system the amplification factor has no dependence on the $q$ coordinate. The previous equation can be rewritten as:

\begin{equation}
 F(t) 
= \int_{-\infty}^\infty  \mu_t(p) S_{1D}\left(p-p_s(t)\right) \mathrm{d}p,
\label{eqn:ft}
\end{equation}
\\
where we have defined the one dimensional flux function as:
\begin{equation}
 S_{1D}(p-p_s(t)) = \int_{-\infty}^\infty S_{2D}(p-p_s(t), q-q_s(t)) \mathrm{d}q
\end{equation}

This representation is a valid approximation only when the apparent size of the source is much smaller than the corresponding Einstein angle of the lens. In this context 
all the information about the source shape and brightness that can be contained in the lightcurve is exhaustively given by the 1D flux function.
In other words, if two sources with different $S_{2D}$ have the same $S_{1D}$ they cannot be distinguished by studying their lightcurves.

\section{Models for extended sources}

In the present study we are analysing three types of sources with different surface brightness: 
\begin{enumerate}
 \renewcommand{\theenumi}{(\arabic{enumi})}
  \item a rotationally symmetric source with a bivariate gaussian surface brightness distribution,
  \item a disk source with constant surface brightness distribution,
  \item a crescent shaped source with constant surface brightness distribution.
\end{enumerate}
The first two sources are the typical choices used in the literature to describe the luminous parts of a quasar (Prasenjit should give some citations here). The third one is a recently proposed
variant \citep{2013MNRAS.434..765K}.

\subsection{Rotationally symmetric source with a bivariate gaussian surface brightness distribution}

A symmetric 2D gaussian can be described mathematically as:

\begin{equation}
 S_{2D}^G(p-p_s, q-q_s) = \frac{S_0^G}{2 \pi \sigma^2} e^{-\frac{(p-p_s)^2}{2 \sigma^2}} e^{-\frac{(q-q_s)^2}{2 \sigma^2}}.
\end{equation}
\\
The corresponding 1D brightness is:

\begin{equation}
 S_{1D}^G(p-p_s) = \frac{S_0^G}{\sqrt{2 \pi} \sigma} e^{-\frac{(p-p_s)^2}{2 \sigma^2}}.
\end{equation}
\\
Other parameters of the model are the total flux $S_0^G$ and $\sigma$.

Although such a definition for a source would have non-zero surface/linear brightness for any coordinate $p,q$, the amount of light received by a detector from outside a $3 \sigma$ disk centered at $p_s, q_s$ 
would be insignificant. For a gaussian distributed surface brightness source the half-light radius is directly proportional to the parameter $\sigma$ according to the equation:
\begin{equation}
r_{1/2} = \sqrt{ln(4)} \sigma.
\end{equation}

\subsection{Disk source with constant surface brightness distribution}

One can construct mathematically a disk source with constant surface brightness and radius $R$ using a stepfunction:
\begin{equation}
 S_{2D}^D(p-p_s, q-q_s) = \frac{S_0^D}{\pi R^2} \Theta \left( R^2 - \left( p-p_s \right)^2 - \left( q-q_s \right)^2 \right).
\end{equation}
By integrating over $q$ coordinate the linear brightness function is obtained:

\begin{equation}
 S_{1D}^D(p-p_s) = \frac{2 S_0^D}{\pi R}  \sqrt{1 - \frac{(p-p_s)^2}{R^2} }    \Theta \left( R^2 - \left( p-p_s \right)^2 \right).
\end{equation}
\\
The half-light radius of a uniform disk source is $R/\sqrt{2}$.

\subsection{Crescent source with constant surface brightness distribution}\label{subsec:crescent}

The surface brightness distribution of a geometric crescent can be built by considering two disk sources of constant brightness. One larger disk will contribute positively to the total flux, while one smaller disk 
that is interior to the large one will contribute negatively. This superposition can be written for 2D as:\\

\begin{equation}
 S_{2D}^C =  S_{2D}^{Dp} -  S_{2D}^{Dn}   (n+1)
 \label{eqn:s2d}
\end{equation}
with\\

\begin{equation}
 S_{2D}^{Dp}(p-p_{sn}, q-q_{sn}) = \frac{S_0^{Dp}}{\pi R_p^2} \Theta \left( R_p^2 - \left( p-p_{sp} \right)^2 - \left( q-q_{sp} \right)^2 \right)
\end{equation}
\\
and
\begin{equation}
 S_{2D}^{Dn}(p-p_{sn}, q-q_{sn}) = \frac{S_0^{Dn}}{\pi R_n^2} \Theta \left( R_n^2 - \left( p-p_{sn} \right)^2 - \left( q-q_{sn} \right)^2 \right).
\end{equation}
\\
The following notations were used: $R_p, (p_{sp}, q_{sp}), R_n, (p_{sn},q_{sn})$ are the radii and coordinate of the center for the larger positive disk and smaller negative disk respectively.  $S_0^{Dp},S_0^{Dn}$ represent the total flux of radiation received from the large and small disk. From this point forward we will not use the total flux from each source. Instead we will 
use the difference which in this case is th total flux from the crescent-shaped source $S_0^C$. \\
Equation \ref{eqn:s2d} can be written as:\\

\begin{align}
 S_{2D}^C &= \frac{S_0^C}{\pi \left(R_p^2-R_n^2 \right)} \left\{ \Theta \left[ R_p^2 - \left( p-p_{sp} \right)^2 - \left( q-q_{sp} \right)^2 \right] \right.\nonumber\\
 &\qquad \left. {} -  \Theta \left[ R_n^2 - \left( p-p_{sn} \right)^2 - \left( q-q_{sn} \right)^2 \right] \right\}.
\end{align}
\\
Analogous for the linear brightness function:

\begin{align}
 S_{1D}^C &= \frac{2 S_0^C}{\pi \left(R_p^2-R_n^2 \right)} \left\{ \sqrt{R_p^2 - (p-p_{sp})^2}  \Theta \left[ R_p^2 - \left( p-p_{sp} \right)^2 \right] \right.\nonumber\\
 &\qquad \left. {} - \sqrt{R_n^2 - (p-p_{sn})^2 } \Theta \left[ R_n^2 - \left( p-p_{sn} \right)^2 \right] \right\}.
\label{eqn:s1_d}
\end{align}

There are some constraints on the parameters used to define a crescent in the previously presented manner that need to be stated. First, we must impose the obvious $R_p > R_n$ relation. Secondly, 
the small disk must always be interior to the large disk:
\begin{equation}
 R_p \ge R_n + \sqrt{\left(p_{sp} - p_{sn} \right)^2 + \left(q_{sp} - q_{sn} \right)^2}
\end{equation}

For the distances between the centers of the two disks we will use the same notations as the one found in the paper \citep{2013MNRAS.434..765K}, $a \equiv p_{sn} - p_{sp}$ and $b \equiv q_{sn} - q_{sp}$.

The half-light radius of any source is invariant to any rotational transformation. In the present case of a crescent source the effective radius is dependent on the parameters $R_p$, $R_n$ and $\sqrt{a^2+b^2}\equiv c $ exclusively. From symmetry considerations the centroid of the source is collinear with the centers of the two disks and it is situated at a distance $d_c$ from the center of the bright disk. $d_c$ can be computed numerically with the use of a variation of equation \ref{eqn:s1_d}:

\begin{equation}
\frac{S_0^C}{2} =  \frac{2 S_0^C}{\pi \left(R_p^2-R_n^2 \right)} \int_{d_c}^{R_p} \left[ \sqrt{R_p^2 - p^2} - \sqrt{R_p^2 - \left(p-c\right)^2} \Theta \left(R_p^2 - \left(p-c\right)^2 \right) \right] dp.
\end{equation}

With the position of the centroid determined, the half-light radius can be also be computed numerically:

\begin{equation}
\begin{aligned}
S_0^C &=  \frac{4S_0^C}{\pi \left(R_p^2-R_n^2 \right)} \int_{0}^{r_{1/2}} \left[ \Theta_2  - \Theta_1  \right] p dp; \\
\Theta_1 &= \bigg[ \pi - \arccos \frac{R_p^2 + p^2 -d_c^2}{2 R_p p} \\
         & \phantom{= \bigg[ \pi} - \arccos \frac{R_p^2 - p^2 + d_c^2}{2 R_p d_c}
            \bigg] \Theta \left( p - R_p + d_c \right); \\
\Theta_2 &=  \pi - \arccos \frac{ p^2  + (c+d_c)^2 - R_n^2}
                                {2 p \left(c + d_c \right)} \\
         & \phantom{= \pi - } \Theta \left( p + R_n - d_c -c \right). 
\end{aligned}
\end{equation}

\begin{figure}
\includegraphics[width = .49\textwidth]{plots/figure_rhalf.eps}
\caption{\label{fig:geom_crescent} Geometry of crescent sources.}
\end{figure}

\section{Lightcurves of the extended sources during fold crossing}

Using equation \ref{eqn:ft} and the one-dimensional flux function presented in the previous section one can compute numerically the lightcurves of the three extented sources for the simplified 
infinite-wall-caustics model.

\subsection{Lightcurve of the gaussian source}

The amount of light received by an observer from a source with a gaussian distributed brightness with $\sigma$ and total flux $S_0^G$ in the absence of any gravitational lensing is:
\begin{equation}
 F^G(t) = \int_{-\infty}^\infty  \left( \mu_0 + \frac{C_0}{\sqrt{p}} \Theta \left( p \right) \right) \left( \frac{S_0^G}{\sqrt{2 \pi} \sigma} e^{-\frac{(p-p_s(t))^2}{2 \sigma^2}} \right) \mathrm{d}p.
\end{equation}
which can be simplified to:
\begin{equation}
 F^G(t) = \mu_0 S_0^G + \frac{C_0 S_0^G}{\sqrt{2\pi} \sigma} \int_{0}^\infty \frac{e^{-\frac{(p-p_s(t))^2}{2 \sigma^2}}}{\sqrt{p}} \mathrm{d}p.
\end{equation}

\begin{figure}
\centering
	\includegraphics[width = 0.48\textwidth]{plots/4source_profile.eps}
	\includegraphics[width = 0.48\textwidth]{plots/4source_magnification.eps}
\caption{\label{fig:lightcurve_gauss} Lightcurves of crescents, gaussian and disk for sources with identical $S_0$ and $r_{1/2}$. Crescent source has $r_{1/2}$ = 0.72754 $R_p$, $R_n$ = 0.4 $R_p$, a = 0.3/-0.3 $R_p$ }
\end{figure}

\subsection{Lightcurve of the disk shaped source}

Analogous to the gaussian shaped source, the disk source with uniform brightness, radius $R$ and unmagnified flux $S_0^D$ has a lightcurve described by the equation:

\begin{equation}
 F^D(t) = \int_{-\infty}^\infty  \left( \mu_0 + \frac{C_0}{\sqrt{p}} \Theta \left( p \right) \right) \left( \frac{2 S_0^D}{ \pi R} \sqrt{1 - \frac{\left( p-p_s(t) \right)^2}{R^2}} \Theta \left(R^2 - \left(p-p_s(t) \right)^2 \right) \right) \mathrm{d}p.
\end{equation}

which is equivalent to:
\begin{equation}
 F^D(t) = \mu_0 S_0^D + \frac{2 C_0 S_0^D}{\pi R} \int_{max(0, p_s(t) - R)}^{max(0, p_s(t) + R)} \frac{1}{\sqrt{p}} \sqrt{1 - \frac{\left( p-p_s(t) \right)^2}{R^2}} \mathrm{d}p.
\end{equation}

\begin{figure}
\centering
	\includegraphics[width = 0.48\textwidth]{plots/4avar_profile.eps}
	\includegraphics[width = 0.48\textwidth]{plots/4avar_magnification.eps}
\caption{\label{fig:lightcurve_disk} Lightcurves of crescent sources with $R_n$ = 0.4 $R_p$.}
\end{figure}

\subsection{Lightcurve of the crescent shaped source}

The lightcurve of a crescent shaped source with unamplified flux $S_0^C$, radii $R_p$, $R_n$ and center displacement $a(t)$ is:
\begin{equation}
 F^c(t) = \mu_0 S_0^C + C_0 \frac{2 S_0^C}{\pi \left( R_p^2 -R_n^2 \right) } 
\left(\int_{max(0, p_s(t) - R)}^{max(0, p_s(t) + R)} \sqrt{\frac{R_p^2 - \left( p-p_s(t) \right)^2 }{p}} \mathrm{d}p 
  -  \int_{max(0, p_s(t) - a(t) - R)}^{max(0, p_s(t) -a(t) + R)} \sqrt{\frac{R_p^2 - \left( p-p_s(t) +a(t) \right)^2 }{p}}  \mathrm{d}p \right)
\end{equation}

\begin{figure}
\centering
	\includegraphics[width = 0.48\textwidth]{plots/4rn_profile.eps}
	\includegraphics[width = 0.48\textwidth]{plots/4rn_magnification.eps}
\caption{\label{fig:lightcurve_crescent_back} Lightcurves of backwards crescents with  a = 0.3 $R_p$.}
\end{figure}

The function $p_s(t)$ can be chosen to be equal to $v_p(t-t_0) + p_{s0}$. Where $p_{s0}$ is the coordinate $p$ of the source at the initial time, and $v_p$ is the component of the velocity
along the $p$ axis. Such a modelling of the motion of the object in the source plane describes a linear motion with constant velocity. Furthermore, we reduce the complexity of the model by 
choosing the function $a(t)$ to be constant in time.

\begin{figure}
\centering
	\includegraphics[width = 0.48\textwidth]{plots/4rn2_profile.eps}
	\includegraphics[width = 0.48\textwidth]{plots/4rn2_magnification.eps}
\caption{\label{fig:lightcurve_crescent} Lightcurve of forwards crescents with a = -0.3 $R_p$.}
\end{figure}

There are four characteristic points visible on the resulting one dimensional light profile of the crescent shaped source. Two of the points, p1 and p4, mark the outer boundaries of the luminous disc component. The other two, p2 and p3, mark the boundaries of the dark disc component. Since $S_1D$ is a projection of $S_2D$ on a line perpendicular to the caustic, the following relation holds:
\begin{equation}
	p_2-p_1 = R_p -R_n - a.
\end{equation}
In addition there are two other obvious relations the characteristic points which are independent of the projection:
\begin{equation}
	p_4 -p_1 = 2 R_p,
\end{equation}

\begin{equation}
        p_3 -p_2 = 2 R_n.
\end{equation}
All four points mark the positions where derivative $\frac{dS_{1D}}{dp}$ is discontinuous. The points can be used to define three regions: $p_1 - p_2$ where $S_1D$ is convex, $p_2 - p_3$ where $S_1D$ is concave, and $p_3 - p_4$ where $S_1D$ is convex again. \\

Due to the nature of the caustic and the monotonic behaviour of the magnification map on both sides of the caustic. The previously mentioned characteristic points are inherited by the microlensing lightcurve. The points on the temporal dimension $t_1, t_2, t_3$ and $t_4$ correspond to instances in time when the fold is aligned with $p_1, p_2, p_3$ and $p_4$, respectively. For a constant relative velocity $v_p$ between the source and the caustic there is a simple relation between the points $p_i$ and instances $t_i$:

\begin{equation}
	t_j - t_i = \frac{p_j - p_i}{v_p}.
\end{equation}

With the use of the previous four equations the following identities can be written:

\begin{equation}
	R_p = \frac{v_p \left( t_4 -t_1 \right)}{2}, 
\end{equation}

\begin{equation}
        R_n = \frac{v_p \left( t_3 -t_2 \right)}{2}, 
\end{equation}

\begin{equation}
        a = \frac{v_p \left( t_4 +t_1 - t_3 - t_2 \right)}{2}. 
\end{equation}

\begin{figure}
\centering
	\includegraphics[width = 0.48\textwidth]{plots/ch_points.eps}
        \includegraphics[width = 0.48\textwidth]{plots/ch_instances.eps}
\caption{\label{fig:char_points} The characteristic points and instances of a crescent source's one dimensional profile and microlensing lightcurve. }
\end{figure}

Figure 4 reveals that the lighcurve of a crescent source has more visible features than the other two light-curves
corresponding to the disc and guasssian shape. There are three regions where $S_{1D}$ and by inheritance the lighcurve
 have distinguishable behaviour. The first region that would be recorded on a lightcurve plot represents the period
of time when the bright disk begins to be overlapped by the caustic and stops when just before the dark disk reaches
the caustic. During this period of time the flux of light from the source is increasingly magnified. The second
period starts and ends with the overlapping of the dark disc. As a boundary of the two regions there is a
distinguishable point where slope of the magnification is drastically changed. This apparent discontinuity in the
first derivative of the magnification function is caused by the caustic amplification of the sudden drop in the $S_{1d}$
function. During the respective period the magnification growth slows or even reverses during the first part of the
period and starts to grow faster again as the dark disk ends its overlap with the caustic. At the point where
the dark disc clears the caustic the growth of the magnification is infinite, which appears as a saddle point
on the lightcurve. Next, the final period correspond to the case when the dark disc has cleared the caustic
and the bright disc continues to overlap with the caustic. During this period the lighcurve reaches a peak
that for most of the parameter space is global and for the rest of the parameter space local.
At the end of the period the growth of the magnification is negative infinite. the respective point appears 
as a second saddle point on the lightcurve. Past this point the magnification of any finite source will 
decay in roughly the same manner. \\

The impact of parameter $a$ on the shape of the lightcurve can be observed in figure 5 (a-var). Sources where 
the center of the dark disc reaches the caustic before the center of the bright disc are characterized by a smoother
broad peak in contrast to the cases where the center of the dark disc reaches the caustic after the bright disk.
In the case of the latter the instance when the dark disc reaches the caustic corresponds to larger and
larger magnifications until it becomes a local and even global peak. The effect of the $R_n$  parameter on the
lighcurve is presented in figures 6 and 7. For the particular set of parameters where $R_p = R_n +a$ the third
period of time discussed previously does not exist. Particularizing further, if the value of the radius of the dark
disc is comparable to the value of the radius of the bright one then the position and shape of the maximum magnification
 are strikingly different. In case the crescent reaches the caustic with the bright region first, the peak magnification
 happens when the dark region reaches the caustic and it is characterized by a sharp variation in magnification growth,
In the opposite case the peak appears before the end of the bright disc reaches the caustic and shape is smoother. \\

\subsection{Microlensing a simulated image of M87}

\begin{figure}
\centering
        \includegraphics[width = 0.48\textwidth]{plots/M87_image.eps}
\caption{\label{fig:M87_image} The simulated image of M87's black hole sillouette obtained by \citep{2012MNRAS.421.1517D} from a GRMHD simulation presented in \citep{2009MNRAS.394L.126M}. The color bar is linear and scaled to reach value 1 for the maximum brightness. Points p2 and p3, mark the start and end of the equivalent dark disk and are associated to the moment of their overlap with the caustic t2 and t3.}
\end{figure}

\citep{2012MNRAS.421.1517D} have created a radiative image of M87 based on the GRMHD simulations presented in \citep{2009MNRAS.394L.126M}. The respective image, reproduced in figure \ref{fig:M87_image}, has been projected to a 1D profile associated to the perpendicular direction to a fold caustic approaching the image from the right. The projection is presented in the left panel of figure 10. The amplification values of the flux of light corresponding to a microlensing event are presented in the right panel of figure 10. In general the behaviour of the lightcurve is similar to the geometric crescent source with the caveat that the outer regions surrounding the luminous parts of the image have non-zero flux and thus are more extended than the simplified source model.

\begin{figure}
\centering
	\includegraphics[width = 0.48\textwidth]{plots/M87_shape.eps}
        \includegraphics[width = 0.48\textwidth]{plots/M87_lc.eps}
\caption{\label{fig:M87_plots} The one dimensional profile obtained from the numerical integration along the ordonata axis of the source image (figure 9) is presented in the left subplot. The corresponding lightcurve associated to the 1D profile is presented in the right subplot. The points p2 and p3 mark the start and end of the equivalent dark disk and are associated to the moments of their overlap with the caustic: t2 and t3.}
\end{figure}