### 2016-10-10 paper1/ms.tex
changes

\documentclass[usenatbib]{mn2e}

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{multicol}

\title{Microlensing as a possible probe of event-horizon structure in quasars}
\author[Tomozeiu et al]{Mihai Tomozeiu\thanks{mihai@physik.uzh.ch},$^{1,2}$ 
Irshad Mohammed,$^{1,2,3}$ 
Manuel Rabold,$^{1,2}$
Prasenjit Saha,$^{1,2}$  
\newauthor
and Joachim Wambsganss$^{1,4}$\\
$^1${Physik-Institut, University of Zurich, Winterthurerstrasse 190,
  8057 Zurich, Switzerland} \\
$^2${Institute for Computational Science, University of Zurich,
  Winterthurerstrasse 190, 8057 Zurich, Switzerland} \\
$^3${Theoretical Astrophysics Group, Fermi National Accelerator Laboratory, Batavia, IL 60510, USA}\\
$^4${Zentrum f\"ur Astronomie der Universit\"at Heidelberg,
  M\"onchhofstrasse 12--14, 69120 Heidelberg, Germany}
}

\begin{document}

\maketitle

\begin{abstract}

In quasars which are lensed by galaxies, the point-like images sometimes show sharp and uncorrelated brightness 
variations (microlensing). These brightness changes are associated with the innermost region of the quasar passing 
through a complicated pattern of caustics produced by the stars in the lensing galaxy. In this paper, we study 
whether the universal properties of optical caustics could enable extraction of shape information about the central 
engine of quasars. We present a toy model with a crescent-shaped source crossing a fold caustic. The silhouette 
of a black hole over an accretion disk tends to produce roughly crescent sources. When a crescent-shaped source 
crosses a fold caustic, the resulting light curve is noticeably different from the case of a circular luminosity 
profile or Gaussian source. With good enough monitoring data, the crescent parameters, apart from one degeneracy, 
can be recovered.

\end{abstract}

\begin{keywords}
Supermassive black holes, microlensing, quasars.
\end{keywords}

\section{The General Picture }

Active galactic nuclei (AGN) are thought to be powered by the accretion of
matter from the proximal environment into a supermassive black hole.
The radiation emitted excites the surrounding medium which becomes
detectable as narrow line regions, broad line regions and optical
continuum.  In the direction perpendicular to the accretion disc,
where the medium is more transparent, the outflow tends to be
collimated to jets.  If the orientation is such that the observer's
view of the central region is not blocked by the accretion disc or by
a jet, a quasar is seen \citep[e.g.,][]{1984RvMP...56..255B}.  While
the basic mechanism \citep[originating in the work
  of][]{1964ApJ...140..796S,1964SPhD....9..246Z,1969Natur.223..690L}
is not in doubt, the central engines, near the event horizons of the
black holes, remain to be probed.

Modestly active galactic nucleii are present in the Galactic centre
and in M87.  In these two cases, the central engines ($<0.1\rm\,mas$
on the sky) are close to being resolved through very long baseline
interferometry, which shows preliminary indications of the
jet-launching structures
\citep{2008JPhCS.131a2055D,2012Sci...338..355D,2016ApJ...820...90F}.
Current data do not deliver images but require fitting to predefined
models for the images.  A whole range of models have been applied,
starting from simple geometric models to more complex physical models
\citep{2008Natur.455...78D,2011ApJ...738...38B,2009ApJ...706..497M,2010ApJ...717.1092D}.
The more complicated models nonetheless tend to predict a crescent
shaped silhouette of the black hole.  This motivated
\cite{2013MNRAS.434..765K} to use a simple geometric crescent model to
fit the available sub-mm interferometric data, They argued that the
crescent is nothing but the silhouette of the event horizon.
The great majority of quasars, however, lie at redshifts beyond 2
\citep{2014A&A...563A..54P} and their central engines would be orders
of magnitude smaller on the sky. The direct observations of the black
hole silhouettes of quasars are far beyond foreseeable
instrumentation.

In the present paper, we consider a possible indirect method, related
to \cite{1999ApJ...524...49A} and \cite{2015ApJ...814L..26M}, 
through which gravitational
microlensing could probe the black hole shadow and its proximal quasar
environment.

Gravitational microlensing of quasars, reviewed in
Section~\ref{sec:microlensing} below, refers to sharp changes in the
observed brightness of quasars that have been lensed by an intervening
galaxies, without any changes in the intrinsic luminosity.  It is a
consequence of two things: the very small size of the central engine,
and granularity of the mass distribution of a lensing galaxy due to
stars.  The latter means that the local magnification is not a smooth
function of source position. It contains a complicated network of
singular curves, known as caustics.
Figure~\ref{fig:magnification_map} shows part of a magnification map
with a few caustics.  The lensed brightness would be given by placing
the source on such a magnification map and integrating the surface
brightness weighted by the magnification.  Most astrophysical sources
straddle several caustics, and hence, their net brightness varies
smoothly with location.  The central engine of quasars, however, is
smaller than the typical spacing between caustics.  As a result, the
lensed brightness undergoes sudden changes as a quasar crosses a
caustic.  Quasar microlensing is mostly observed in the optical
continuum \citep[e.g.,][]{2012A&A...544A..62S} but is also known in
x-rays \citep[e.g.,][]{2016arXiv160207601N} and has been argued for at
cm wavelengths \citep{2000A&A...358..793K}.  The effect supplies an
upper limit on the size of the central engine
\citep[e.g.,][]{2015ApJ...814L..26M} and can be used to study the mass
distribution and kinematics of stars in the lensing galaxy as well
\citep[e.g.,][]{2012ApJ...744..111P}.  Microlensing caustics have an
additional remarkable feature: though they can be very complicated,
they have some universal properties well-known from catastrophe
theory.  In particular, very close to the simplest caustics (known as
folds), the magnification is approximately constant on one side and
$\propto1/\sqrt p$ where $p$ is the transverse distance of the source
from the caustic.  This property will be exploited later.

In Section~\ref{sec:source-models} we introduce the three source
profiles used in our subsequent models and simulations: a
constant-brightness disc, a circular Gaussian, and the crescent source
introduced by \citep{2013MNRAS.434..765K}.  The latter is simply a
constant-brightness disc with a smaller, non-concentric disc cut out
of it.  We also derive the half-light radius for a crescent.  The
half-light radius can characterise the source size for all three types
of source.

Section~\ref{sec:fold-crossing} shows the light curves that result
when each of the model sources crosses an ideal fold.  This would
apply in Figure~\ref{fig:magnification_map} to sources along the path
AB or BC, for sources small enough that the curvature of the caustics
is negligible.  With this assumption one can imagine the caustic as an
infinite wall to be crossed by the source as presented in
Figure~\ref{fig:infinite_fold}.  The source brightness distribution
parallel to the caustic naturally makes no difference to the
observable brightness; each source can be replaced by an effective
one-dimensional source profile, by flattening the source so it becomes
perpendicular to the caustic.  In principle, the effective
one-dimensional brightness profile could be recovered from the light
curve by deconvolution.  \cite{1999ApJ...524...49A} modelled this
profile as the result of a circular accretion disc seen through the
spacetime around a Kerr black hole, and \cite{2012MNRAS.423..676A}
have applied the idea to observed light curves to infer properties of
quasar accretion discs.  A similar idea appears in
\cite{2013ApJ...769..131C}.  In this work, we take a simpler but
arguably more robust approach: we study features in the light curves
characteristic of a crescent-like source which in turn would indicate
a black-hole silhouette.  Figure~\ref{fig:char_points} shows the
qualitative features: there is a period during which the dark cutout
disc is crossing the caustic, and before and after there are periods
when the only the bright parts of the crescent are in transit across
the caustic.  The details depend on the orientation of the crescent,
but basically the dark disc causes a rising light curve to plateau or
dip.  These features are still present, albeit faintly, if the simple
crescent is replaced by a source based on an accretion-disc simulation
of a black-hole environment (Figure~\ref{fig:M87_plots}).

In Section~\ref{sec:numerics} we carry out source fitting to
lightcurves, with both noise and systematic errors are present.  We
generate three lightcurves by taking a uniform disc, a circular
Gaussian, and a crescent across the path AB in
Figure~\ref{fig:magnification_map}, and then adding noise.  The path
simulates crossing a clean but not ideal fold.  In addition, we
generate templates lightcurves by running the three source types, with
various parameter values, across the path CB.  That is, the templates
come from a similar but not identical caustic, thus deliberately
generating a systematic error.  We then fit the noisy lightcurves to
the templates using Markov chain Monte-Carlo.  We find that the
correct source type can be inferred from the $\chi^2$ values.  The
parameter values can also be inferred. The fitting errors are larger
than the formal uncertainties, which is expected in the presence of
systematic errors, but still appear acceptable.

Finally in Section~\ref{sec:discussion} we discuss in more depth the
implications of the results presented in the previously mentioned
sections.  One of the most interesting implication is the possibility
to estimate the black hole mass from the reconstructed
parameters. This further requires good approximations of the relative
transversal velocity between the quasar and the lens. Proper
constraints can be set with independent observations of the stellar
structure that contain the gravitational lens.

The thorough study of the possibility to reconstruct the quasar's
structural parameters from light-curves containing multiple
microlensing events represents the target of future work. This will
most probably require the use of powerful statistical tools. Another
path for future work is the designing of an observation regime best
suited for acquiring the necessary microlensed light-curve.

\section{Microlensing}\label{sec:microlensing}

We start by introducing in a succinct manner the gravitational lensing
theory that is relevant for microlensing in general and for the scope
of the present paper in particular.  More detailed presentation of the
theory can be found in several references, such as
\cite{2001stgl.book.....P}.

\subsection{Magnification}

The gravitational lens equation
\begin{equation}
\vec\beta = \vec\theta - \vec\alpha(\vec\theta)
\label{eqn:lens}
\end{equation}
relates the apparent sky position $\vec\theta$ of a light source to
its true but unobservable sky position $\vec\beta$ through the bending
angle $\vec\alpha$.  The latter is an integral 
\begin{equation}
\vec\alpha = (1+z_{_L})\frac{D_S-D_L}{D_SD_L} \frac{4G}{c^2}
\int \Sigma(\vec\theta')
\frac{\vec\theta-\vec\theta'}{|\vec\theta-\vec\theta'|^2}\,
d^2\vec\theta'
\label{eqn:alpha}
\end{equation}
depending on the projected density $\Sigma(\vec\theta)$ (kg/steradian)
of the lens, the lens redshift $z_{_L}$ and the comoving distances
$D_L$ and $D_S$ to the lens and source.  The derivative of the
apparent position with respect to the source position
\begin{equation}
M(\vec\theta) =
\left(\frac{\partial\beta}{\partial\theta}\right)^{-1}
\label{eqn:magnif-matrix}
\end{equation}
is known as the magnification matrix, and its determinant
\begin{equation}
\mu(\vec\theta) = \det|M(\vec\theta)|
\end{equation}
is the brightness amplification of an image of a point source.  In
other words, the source will brighten or dim according to whether
$\mu(\vec\theta)$ is more or less than unity.  If there are multiple
mages at distinct but not observationally resolved $\vec\theta_i$ from
the same $\vec\beta$, a total brightness amplification
of\begin{equation} \mu_{\rm total} = \sum_{i} \mu(\vec\theta_i)
\end{equation}
applies. It is possible for the magnification to become formally
infinite, as a result of an eigenvalues of the magnification matrix
(\ref{eqn:magnif-matrix}) becoming infinite.  This generically happens
on curves on the $\vec\theta$ plane, known as {\em critical curves}.
Mapping a critical curve to the source plane, through the lens
equation (\ref{eqn:lens}), to the source plane $\vec\beta$ gives the
so-called caustics.  Caustics can appear in the optical system, not just
gravitational lensing.  For a point source, caustics are singularities
of the magnification; for finite-size sources caustics correspond to
high and sharply changing magnification.

Caustics are important in all forms of lensing with multiple images,
but they have a special significance for lens quasars, first pointed
out by \cite{1979Natur.282..561C}.  The granularity of the mass
distribution $\Sigma(\vec\theta)$ due to individual stars produces a
caustic network on the scale of $\sqrt{GM_\odot D_L/c^2}$ or $10^{-6}
\sqrt{M/M_\odot}\rm\,arcsec$ for typical lens and source redshift
\citep{2001PASA...18..207W}.  Extended sources wash out this
micro-caustic structure, but the optical-continuum source of quasars
is even smaller.

\begin{figure}
\centering
\includegraphics[width=0.9\hsize]{figures/IRIS567_path.eps}
\caption{\label{fig:magnification_map} Magnification map used later in
  this paper (Section~\ref{sec:numerics}). The white line marks the
  trajectory of the center of the sources for which the lightcurves
  presented in figure 11 are generated.  Note that this represents an
  atypically simple region of any realistic magnification map.}
\end{figure}

\subsection{Magnification near a fold caustic}

A simple example of a caustic network is shown in
Figure~\ref{fig:magnification_map}.  There are two general categories
of caustics in gravitational lensing, cusps and fold, and examples of
both kinds can be seen in this figure.  Magnification near a caustic
has universal properties, independent of the system and has been
extensively studied
\citep{1986ApJ...310..568B,1992A&A...260....1S,2002ApJ...574..970G,2002ApJ...580..468G}.
In particular, at distance $p$ from a fold caustic
\begin {equation}
 \mu(p) = \mu_0 + C_0 \frac{1}{\sqrt{p}} \Theta(p).
\end {equation}
Here the magnification of a point source near a caustic is equal to
the sum of the magnification due to other reasons $\mu_0$ 
(such as zero point magnification and the effect of other weaker and 
distant lenses), assumed to
be locally constant, and a decrease with the square root of the
distance from the fold. The latter term becomes activated only after
the source enters the region interior to the caustic curves when the
values of the step function $\Theta(p)$ become unity. The
proportionality constant $C_0$ depends on the local conditions in the
vicinity of the caustic.

A source of arbitrary shape can be described by a two-dimensional brightness function $S_{2D}(p - p_s, q - q_s)$ defined for a coordinate system $p,q$ where $p_s, q_s$ denote the coordinates of the center of a source.

\begin{figure}
\includegraphics[width = .49\textwidth]{figures/infinite_fold_ar.eps}
\caption{\label{fig:infinite_fold} Source profiles for Gaussian, disk,
  crescents (right) and magnification map for an infinite fold
  (left). Objects have the same $S_{0}$ and $r_{1/2}$. The black arrow marks the direction of motion of the
sources relative to the caustic.}
\end{figure}

For a microlensing event the lightcurve can be written for an
undefined source shape as:
\begin{equation}
 F(t) = \int_{-\infty}^\infty \int_{-\infty}^\infty S_{2D}(p-p_s(t), q-q_s(t)) \mu_t(p) \mathrm{d}q \mathrm{d}p
 \label{eqn:ft2d}
\end{equation}
In order to build the previous equation we have considered that the
time dependency of the flux $F$ is given only by the motion of the
source with respect to a fixed caustic. Therefore the only time
dependent quantities in the right hand side of the equation are the
coordinates of the center of the source $p_s,q_s$ and by construct
$S_{2D}$.

Due to the choice of the coordinate system the amplification factor has no dependence on the $q$ coordinate. The previous equation can be rewritten as:

\begin{equation}
 F(t) 
= \int_{-\infty}^\infty  \mu_t(p) S_{1D}\left(p-p_s(t)\right) \mathrm{d}p,
\label{eqn:ft}
\end{equation}
\\
where we have defined the one dimensional flux function as:
\begin{equation}
 S_{1D}(p-p_s(t)) = \int_{-\infty}^\infty S_{2D}(p-p_s(t), q-q_s(t)) \mathrm{d}q
\end{equation}

This representation is a valid approximation only when the apparent size of the source is much smaller than the corresponding Einstein angle of the lens. In this context 
all the information about the source shape and brightness that can be contained in the lightcurve is exhaustively given by the 1D flux function.
In other words, if two sources with different $S_{2D}$ have the same $S_{1D}$ they cannot be distinguished by studying their lightcurves.

\section{Models for extended sources}\label{sec:source-models}

In the present study we are analysing three types of sources with different surface brightness: 
\begin{enumerate}
 \renewcommand{\theenumi}{(\arabic{enumi})}
  \item a rotationally symmetric source with a bivariate gaussian surface brightness distribution,
  \item a disk source with constant surface brightness distribution,
  \item a crescent shaped source with constant surface brightness distribution.
\end{enumerate}
The first two sources are the typical choices used in the literature to describe the luminous parts of a quasar (Prasenjit should give some citations here). The third one is a recently proposed
variant \citep{2013MNRAS.434..765K}.

\subsection{Rotationally symmetric source with a bivariate gaussian surface brightness distribution}\label{subsec:gaussian}

A symmetric 2D gaussian can be described mathematically as:

\begin{equation}
 S_{2D}^G(p-p_s, q-q_s) = \frac{S_0^G}{2 \pi \sigma^2} e^{-\frac{(p-p_s)^2}{2 \sigma^2}} e^{-\frac{(q-q_s)^2}{2 \sigma^2}}.
\end{equation}
\\
The corresponding 1D brightness is:

\begin{equation}
 S_{1D}^G(p-p_s) = \frac{S_0^G}{\sqrt{2 \pi} \sigma} e^{-\frac{(p-p_s)^2}{2 \sigma^2}}.
\end{equation}
\\
Other parameters of the model are the total flux $S_0^G$ and $\sigma$.

Although such a definition for a source would have non-zero surface/linear brightness for any coordinate $p,q$, the amount of light received by a detector from outside a $3 \sigma$ disk centered at $p_s, q_s$ 
would be insignificant. For a gaussian distributed surface brightness source the half-light radius is directly proportional to the parameter $\sigma$ according to the equation:
\begin{equation}
r_{1/2} = \sqrt{ln(4)} \sigma.
\end{equation}

\subsection{Disk source with constant surface brightness distribution}

One can construct mathematically a disk source with constant surface brightness and radius $R$ using a stepfunction:
\begin{equation}
 S_{2D}^D(p-p_s, q-q_s) = \frac{S_0^D}{\pi R^2} \Theta \left( R^2 - \left( p-p_s \right)^2 - \left( q-q_s \right)^2 \right).
\end{equation}
By integrating over $q$ coordinate the linear brightness function is obtained:

\begin{equation}
 S_{1D}^D(p-p_s) = \frac{2 S_0^D}{\pi R}  \sqrt{1 - \frac{(p-p_s)^2}{R^2} }    \Theta \left( R^2 - \left( p-p_s \right)^2 \right).
\end{equation}
\\
The half-light radius of a uniform disk source is $R/\sqrt{2}$.

\subsection{Crescent source with constant surface brightness distribution}\label{subsec:crescent}

The surface brightness distribution of a geometric crescent can be built by considering two disk sources of constant brightness. One larger disk will contribute positively to the total flux while one smaller disk 
that is interior to the large one will contribute negatively. This superposition can be written for 2D as:\\

\begin{equation}
 S_{2D}^C =  S_{2D}^{Dp} -  S_{2D}^{Dn}  
 \label{eqn:s2d}
\end{equation}
with\\

\begin{equation}
 S_{2D}^{Dp}(p-p_{sn}, q-q_{sn}) = \frac{S_0^{Dp}}{\pi R_p^2} \Theta \left( R_p^2 - \left( p-p_{sp} \right)^2 - \left( q-q_{sp} \right)^2 \right)
\end{equation}
\\
and
\begin{equation}
 S_{2D}^{Dn}(p-p_{sn}, q-q_{sn}) = \frac{S_0^{Dn}}{\pi R_n^2} \Theta \left( R_n^2 - \left( p-p_{sn} \right)^2 - \left( q-q_{sn} \right)^2 \right).
\end{equation}
\\
The following notations were used: $R_p, (p_{sp}, q_{sp}), R_n, (p_{sn},q_{sn})$ are the radii and coordinate of the center for the larger positive disk and smaller negative disk respectively.  $S_0^{Dp},S_0^{Dn}$ represent the total flux of radiation received from the large and small disk. From this point forward we will not use the total flux from each source. Instead we will 
use the difference which in this case is th total flux from the crescent-shaped source $S_0^C$. \\
Equation \ref{eqn:s2d} can be written as:\\

\begin{align}
 S_{2D}^C &= \frac{S_0^C}{\pi \left(R_p^2-R_n^2 \right)} \left\{ \Theta \left[ R_p^2 - \left( p-p_{sp} \right)^2 - \left( q-q_{sp} \right)^2 \right] \right.\nonumber\\
 &\qquad \left. {} -  \Theta \left[ R_n^2 - \left( p-p_{sn} \right)^2 - \left( q-q_{sn} \right)^2 \right] \right\}.
\end{align}
\\
Analogous for the linear brightness function:

\begin{align}
 S_{1D}^C &= \frac{2 S_0^C}{\pi \left(R_p^2-R_n^2 \right)} \left\{ \sqrt{R_p^2 - (p-p_{sp})^2}  \Theta \left[ R_p^2 - \left( p-p_{sp} \right)^2 \right] \right.\nonumber\\
 &\qquad \left. {} - \sqrt{R_n^2 - (p-p_{sn})^2 } \Theta \left[ R_n^2 - \left( p-p_{sn} \right)^2 \right] \right\}.
\label{eqn:s1_d}
\end{align}

There are some constraints on the parameters used to define a crescent in the previously presented manner that need to be stated. First, we must impose the obvious $R_p > R_n$ relation. Secondly, 
the small disk must always be interior to the large disk:
\begin{equation}
 R_p \ge R_n + \sqrt{\left(p_{sp} - p_{sn} \right)^2 + \left(q_{sp} - q_{sn} \right)^2}
\end{equation}

For the distances between the centers of the two disks we will use the same notations as the one found in the paper \citep{2013MNRAS.434..765K}, $a \equiv p_{sn} - p_{sp}$ and $b \equiv q_{sn} - q_{sp}$.

The half-light radius of any source is invariant to any rotational transformation. In the present case of a crescent source the effective radius is dependent on the parameters $R_p$, $R_n$ and $\sqrt{a^2+b^2}\equiv c $ exclusively. From symmetry considerations the centroid of the source is collinear with the centers of the two disks and it is situated at a distance $d_c$ from the center of the bright disk. $d_c$ can be computed numerically with the use of a variation of equation \ref{eqn:s1_d}:

\begin{equation}
\begin{aligned}
\frac{S_0^C}{2} & = \frac{2 S_0^C}{\pi \left(R_p^2-R_n^2 \right)} \int_{d_c}^{R_p} \bigg[ \sqrt{R_p^2 - p^2} \\ 
        & - \sqrt{R_p^2 - \left(p-c\right)^2} \Theta \left(R_p^2 - \left(p-c\right)^2 \right) \bigg] dp. 
\end{aligned}
\end{equation}

With the position of the centroid determined, the half-light radius can be also be computed numerically:

\begin{equation}
\begin{aligned}
S_0^C &=  \frac{4S_0^C}{\pi \left(R_p^2-R_n^2 \right)} \int_{0}^{r_{1/2}} \left[ \Theta_2  - \Theta_1  \right] p dp; \\
\Theta_1 &= \bigg[ \pi - \arccos \frac{R_p^2 + p^2 -d_c^2}{2 R_p p} \\
         & \phantom{= \bigg[ \pi} - \arccos \frac{R_p^2 - p^2 + d_c^2}{2 R_p d_c}
            \bigg] \Theta \left( p - R_p + d_c \right); \\
\Theta_2 &=  \pi - \arccos \frac{ p^2  + (c+d_c)^2 - R_n^2}
                                {2 p \left(c + d_c \right)} \\
         & \phantom{= \pi - } \Theta \left( p + R_n - d_c -c \right). 
\end{aligned}
\end{equation}

\begin{figure}
\includegraphics[width = .49\textwidth]{figures/figure_rhalf.eps}
\caption{\label{fig:geom_crescent} Geometry of crescent sources: the source can be seen as a superposition of a smaller dark disk over and inside a larger brighter disk. $R_p$ and $R_n$ mark the radii of the bright and dark disk. $d_c$ and $c$ mark the distance from the center of the bright disk to the centroid and from the dark disk to the centroid.  }
\end{figure}

\section{Lightcurves of the extended sources during fold crossing}\label{sec:fold-crossing}

Using equation \ref{eqn:ft} and the one-dimensional flux function
presented in the previous section one can compute numerically the
lightcurves of the three extended sources for the simplified
infinite-wall-caustics model.

\subsection{Lightcurve of the gaussian source}

The amount of light received by an observer from a source with a gaussian distributed brightness with $\sigma$ and total flux $S_0^G$ in the absence of any gravitational lensing is:
\begin{equation}
 F^G(t) = \int_{-\infty}^\infty  \left( \mu_0 + \frac{C_0}{\sqrt{p}} \Theta \left( p \right) \right) \left( \frac{S_0^G}{\sqrt{2 \pi} \sigma} e^{-\frac{(p-p_s(t))^2}{2 \sigma^2}} \right) \mathrm{d}p.
\end{equation}

which can be simplified to:
\begin{equation}
 F^G(t) = \mu_0 S_0^G + \frac{C_0 S_0^G}{\sqrt{2\pi} \sigma} \int_{0}^\infty \frac{e^{-\frac{(p-p_s(t))^2}{2 \sigma^2}}}{\sqrt{p}} \mathrm{d}p.
\end{equation}

\begin{figure}
\centering
    \includegraphics[width = 0.48\textwidth]{figures/S1D_all.eps}
    \includegraphics[width = 0.48\textwidth]{figures/4source_magnification.eps}
\caption{\label{fig:lightcurve_gauss} 1D profiles (top panel) and simulated lightcurves (bottom) of crescent, Gaussian and disk sources with identical total flux ($S_0$) and half-light radius($r_{1/2}$). Crescent source has $r_{1/2}$ = 0.72754 $R_p$, $R_n$ = 0.4 $R_p$, a = 0.3/-0.3 $R_p$. }
\end{figure}

\subsection{Lightcurve of the disk shaped source}

Analogous to the gaussian shaped source, the disk source with uniform brightness, radius $R$ and unmagnified flux $S_0^D$ has a lightcurve described by the equation:

\begin{equation}
\begin{aligned}
 F^D(t) &= \int_{-\infty}^\infty  \left( \mu_0 + \frac{C_0}{\sqrt{p}} \Theta \left( p \right) \right) \\
    & \bigg[ \frac{2 S_0^D}{ \pi R} \sqrt{1 - \frac{\left( p-p_s(t) \right)^2}{R^2}}  \Theta \left(R^2 - \left(p-p_s(t) \right)^2 \right) \bigg] \mathrm{d}p.
\end{aligned}
\end{equation}

which is equivalent to:
\begin{equation}
 F^D(t) = \mu_0 S_0^D + \frac{2 C_0 S_0^D}{\pi R} \int_{max(0, p_s(t) - R)}^{max(0, p_s(t) + R)} \frac{1}{\sqrt{p}} \sqrt{1 - \frac{\left( p-p_s(t) \right)^2}{R^2}} \mathrm{d}p.
\end{equation}

\begin{figure}
\centering
    \includegraphics[width = 0.48\textwidth]{figures/S1D_var_a.eps}
    \includegraphics[width = 0.48\textwidth]{figures/4avar_magnification.eps}
\caption{\label{fig:a_var} 1D profiles (top panel) and simulated lightcurves (bottom) of crescent sources with identical $R_p$ and $R_n$ ($R_n$ = 0.4 $R_p$) and different disk centers displacement (a).}
\end{figure}

\subsection{Lightcurve of the crescent shaped source}

The lightcurve of a crescent shaped source with unamplified flux $S_0^C$, radii $R_p$, $R_n$ and center displacement $a(t)$ is:

\begin{equation}
\begin{aligned}
 F^c(t) &= \mu_0 S_0^C + C_0 \frac{2 S_0^C}{\pi \left( R_p^2 -R_n^2 \right) } \\
    &\bigg[ \int_{max(0, p_s(t) - R)}^{max(0, p_s(t) + R)} \sqrt{\frac{R_p^2 - \left( p-p_s(t) \right)^2 }{p}} \mathrm{d}p \\
    &  -  \int_{max(0, p_s(t) - a(t) - R)}^{max(0, p_s(t) -a(t) + R)} \sqrt{\frac{R_p^2 - \left( p-p_s(t) +a(t) \right)^2 }{p}} \mathrm{d}p  \bigg] .
\end{aligned}
\end{equation}

\begin{figure}
\centering
    \includegraphics[width = 0.48\textwidth]{figures/S1D_var_rn_a_poz.eps}
    \includegraphics[width = 0.48\textwidth]{figures/5Rn_back_var_magnification.eps}
\caption{\label{fig:lightcurve_crescent_back} 1D profiles (top panel) and simulated lightcurves (bottom) of backwards crescent sources with identical $R_p$ and $a$ ($a$ = 0.3 $R_p$) and different dark disk radii ($R_n$).}
\end{figure}

The function $p_s(t)$ can be chosen to be equal to $v_p(t-t_0) + p_{s0}$. Where $p_{s0}$ is the coordinate $p$ of the source at the initial time, and $v_p$ is the component of the velocity
along the $p$ axis. Such a modelling of the motion of the object in the source plane describes a linear motion with constant velocity. Furthermore, we reduce the complexity of the model by 
choosing the function $a$ to be constant in time.

\begin{figure}
\centering
    \includegraphics[width = 0.48\textwidth]{figures/S1D_var_rn_a_neg.eps}
    \includegraphics[width = 0.48\textwidth]{figures/5Rn_forw_var_magnification.eps}
\caption{\label{fig:lightcurve_crescent} 1D profiles (top panel) and simulated lightcurves (bottom) of forwards crescent sources with identical $R_p$ and $a$ ($a$ = - 0.3 $R_p$) and different dark disk radii ($R_n$). }
\end{figure}

In figure \ref{fig:char_points} There are four characteristic points visible on the resulting one-dimensional light profile of the crescent shaped source. Two of the points, p1 and p4, mark the outer boundaries of the luminous disc component. The other two, p2 and p3, mark the boundaries of the dark disc component. Since $S_{1D}$ is a projection of $S_{2D}$ on a line perpendicular to the caustic, the following relation holds:
\begin{equation}
    p_2-p_1 = R_p -R_n - a.
\end{equation}
In addition there are two other obvious relations the characteristic points which are independent of the projection:
\begin{equation}
    p_4 -p_1 = 2 R_p,
\end{equation}

\begin{equation}
        p_3 -p_2 = 2 R_n.
\end{equation}
All four points mark the positions where derivative $\frac{dS_{1D}}{dp}$ is discontinuous. The points can be used to define three regions: $p_1 - p_2$ where $S_1D$ is convex, $p_2 - p_3$ where $S_1D$ is concave, and $p_3 - p_4$ where $S_1D$ is convex again. \\

Due to the nature of the caustic and the monotonic behaviour of the magnification map on both sides of the caustic. The previously mentioned characteristic points are inherited by the microlensing lightcurve. The points on the temporal dimension $t_1, t_2, t_3$ and $t_4$ correspond to instances in time when the fold is aligned with $p_1, p_2, p_3$ and $p_4$, respectively. For a constant relative velocity $v_p$ between the source and the caustic there is a simple relation between the points $p_i$ and instances $t_i$:

\begin{equation}
    t_j - t_i = \frac{p_j - p_i}{v_p}.
\end{equation}

With the use of the previous four equations the following identities can be written:

\begin{equation}
    R_p = \frac{v_p \times  \left( t_4 -t_1 \right)}{2}, 
\end{equation}

\begin{equation}
        R_n = \frac{v_p \times \left( t_3 -t_2 \right)}{2}, 
\end{equation}

\begin{equation}
        a = \frac{v_p \times \left( t_4 +t_1 - t_3 - t_2 \right)}{2}. 
\end{equation}

\begin{figure}
\centering
	\includegraphics[width = 0.48\textwidth]{figures/ch_points.eps}
        \includegraphics[width = 0.48\textwidth]{figures/ch_instances.eps}
\caption{\label{fig:char_points} The characteristic points and instances 
of a crescent source's one dimensional profile and microlensing lightcurve.
In the top panel a typical 1D profile  of a crescent source is presented. 
There are two light gray regions that mark the regions of the bright disk 
that are not overlapped by the dark disk. The dark gray band marks the 
region where the two disks overlap. $p_1$ and $p_4$ mark the start and 
end of the bright disk in the 
1D representation. Analogous, $p_2$ and $p_3$ mark the start and 
end of the dark disk.
The bottom panel represents the associated lightcurve with the points 
$t_i$ marking the moment when point $p_i$ overlaps with the caustic.
The gray shades are analogous to the ones in the top panel.
 }
\end{figure}

Figure \ref{fig:lightcurve_gauss} reveals that the 
lightcurve of a crescent source has more
visible features than the other two light-curves corresponding to the
disc and gaussian shape. There are three regions where $S_{1D}$ and by
inheritance, the lightcurve has distinguishable behaviour. The first
region that would be recorded on a lightcurve plot represents the
period of time when the bright disk begins to be overlapped by the
caustic and stops when just before the dark disk reaches the
caustic. During this period of time, the flux of light from the source
is increasingly magnified. The second period starts and ends with the
overlapping of the dark disc. As a boundary of the two regions, there
is a distinguishable point where the slope of the magnification is
drastically changed. This apparent discontinuity in the first
derivative of the magnification function is caused by the caustic
amplification of the sudden drop in the $S_{1d}$ function. During the
respective period, the magnification growth slows or even reverses
during the first part of the period and starts to grow faster again as
the dark disk ends its overlap with the caustic. At the point where
the dark disc clears the caustic, the growth of the magnification is
infinite, which appears as a saddle point on the lightcurve. Next, the
final period corresponds to the case when the dark disc has cleared
the caustic and the bright disc continues to overlap with the
caustic. During this period, the lightcurve reaches a peak that for
most of the parameter space is global and for the rest of the
parameter space local.  At the end of the period, the growth of the
magnification is negative infinite. the respective point appears as a
second saddle point on the lightcurve. Past this point the
magnification of any finite source will decay in roughly the same
manner.

The impact of parameter $a$ on the shape of the lightcurve can be
observed in Figure~\ref{fig:a_var}. Sources where the center of the
dark disc reaches the caustic before the center of the bright disc are
characterized by a smoother broad peak in contrast to the cases where
the center of the dark disc reaches the caustic after the bright disk.
In the case of the latter the instance when the dark disc reaches the
caustic corresponds to larger and larger magnifications until it
becomes a local and even global peak. The effect of the $R_n$
parameter on the light curve is presented in
Figures~\ref{fig:lightcurve_crescent} and
\ref{fig:lightcurve_crescent_back}. For the particular set of
parameters where $R_p = R_n +a$ the third period of time discussed
previously does not exist. Particularizing further, if the value of
the radius of the dark disc is comparable to the value of the radius
of the bright one then the position and shape of the maximum
magnification are strikingly different. In case the crescent reaches
the caustic with the bright region first, the peak magnification
happens when the dark region reaches the caustic and it is
characterized by a sharp variation in magnification growth, In the
opposite case, the peak appears before the end of the bright disc
reaches the caustic and shape is smoother.

\begin{figure}
\centering
    \includegraphics[width = 0.48\textwidth]{figures/M87_shape.eps}
        \includegraphics[width = 0.48\textwidth]{figures/M87_lc.eps}
\caption{\label{fig:M87_plots} The one dimensional profile obtained
  from the numerical integration along the ordonata axis of a simulated source
  image is presented in the upper panel. The source image was presented in 
 Dexter et al 2012 in the top right-most panel of figure 5 
(of the respective paper) as a result of the simulations 
presented in McKinney \& Blandford 2009. The corresponding
  lightcurve associated to the 1D profile is presented in the lower
  panel. The points $\rm p_2$ and $\rm p_3$ mark the start and 
  end of the equivalent
  dark disk and are associated with the moments of their overlap with
  the caustic: $\rm t_2$ and $\rm t_3$. $\rm h_{w}$ denotes the maximum width of the crescent.}
\end{figure}

\subsection{Microlensing a simulated image of M87}

\cite{2012MNRAS.421.1517D} have created a radiative image of M87 based on the GRMHD simulations presented in \citep{2009MNRAS.394L.126M}. 
The top right-most image in figure 5 of \cite{2012MNRAS.421.1517D} has been projected to a 1D profile associated to the perpendicular 
direction to a fold caustic approaching the image from the right. The projection is presented in the upper panel of figure 9. 
The amplification values of the flux of light corresponding to a microlensing event are presented in the lower panel of figure 9. 
In general, the behaviour of the lightcurve is similar to the geometric crescent source with the caveat that the outer regions 
surrounding the luminous parts of the image have non-zero flux and thus are more extended than the simplified source model.

\section{Fitting mock data}\label{sec:numerics}

Having used the simple model of an ideal fold to gain insight, we now
consider the question of whether the characteristics of a crescent
source could be discerned in real data having noise and systematic
error as well.  Our method is to generate microlensing lightcurves
using a widely-used microlensing code
\citep{1990PhDT.......180W,1999A&A...346L...5W,1999JCoAM.109..353W} to
which we have added a small extension to provide the option of
crescent sources, and then fit these mock lightcurves to source models
using Markov-chain Monte Carlo (MCMC).

\begin{figure}
\centering
  \includegraphics[width=0.5\textwidth]{figures/cc_forward.eps}
\caption{\label{fig:cc_forward} Mock lightcurve fitted with the
  correct source type and caustic.  The upper panel shows the mock
  data points, and running through them is a seven-parameter best fit:
  these being the crescent parameters $(R_p,R_n,a,b)$ and additionally
  the start and end times and the overall normalisation.  (Ideal
  lightcurves for the same crescent parameters appear above in
  Figures~\ref{fig:lightcurve_gauss}, \ref{fig:a_var},
  \ref{fig:lightcurve_crescent} and \ref{fig:char_points}.)  The curve
  has $\chi^2=243, P=0.32$ indicating a good fit.  The lower panel
  shows the posterior distributions for $r_{1/2}$ (solid histogram)
  and $R_n/R_p$ (dashed histogram).}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=0.5\textwidth]{figures/cc_backward.eps}
\caption{\label{fig:cc_backward} Like Figure~\ref{fig:cc_backward},
  except that the fit uses path~II whereas the mock data still come
  from path~I. The curve has $\chi^2=284, P=0.01$ indicating a
  marginally acceptable fit.}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=0.5\textwidth]{figures/gc_forward.eps}
\caption{\label{fig:gc_forward} Like Figure~\ref{fig:cc_backward}, but
  showing the best fit to a Gaussian source.  The histogram shows the
  posterior distribution of $r_{1/2}$ (there is no $R_n$).  It appears
  to show $r_{1/2}$ as accurately recovered, but is not to be trusted,
  because the fit has $\chi^2=344, P=3\times10^{-6}$ indicating
  rejection of the Gaussian source.  Looking closely, the curve is a
  little too shallow near $t=3$, and it is unable to fit the second
  sub-peak near $t=5$.}
\end{figure}

\subsection{Numerical microlensing lightcurves}

The microlensing code uses a ray-shooting technique to compute the
gravitational lensing effect of a mass distribution consisting of
(a)~a smooth component and (b)~a random distribution of point masses
representing stars.  The ray shooting maps a grid of $\vec\theta$ to
$\vec\beta$ using the lens equation (\ref{eqn:lens}).  That is, the
rays are shot from the observer back to the source.  For the
computation of the individual deflection angles
$\vec\alpha(\vec\theta)$ a hierarchical or tree method is used.  The
positions of all lensing masses are put into a grid of $\vec\theta$.
Each grid cell is subdivided into four smaller squares recursively
until every cell contains only one mass.  Nearby masses are added
individually while distance masses are clumped into larger grid cells
whose net contribution is approximated by its first few multipole
moments.  Scaled units are used, with the constant pre-factor in the
deflection angle (\ref{eqn:alpha}) separated out.  The result of ray
shooting is a pixel map on the $\vec\beta$ plane of the number of
lightrays which arrive at the source plane from a particular observer.
This intermediate result is effectively a magnification map on the
source plane.

When generating the magnification map depicted in
Figure~\ref{fig:magnification_map}, which is used in our analysis
below, only two point masses were included.  This was done in order to
have clean fold caustics.  There was no smooth component.

Once the map is created, the lightcurve can be obtained by specifying
the transit path of the source across the map.  The source is moved
along the transit line, one pixel at a time, and at each step the
brightness distribution of the source is convolved
(equation~\ref{eqn:ft2d}) with the magnification map to give the
brightness.  In real life, not only are both lens and source moving
but the lens lens configuration, and with it the magnification
pattern, is also changing with time.  While the first subtlety is
taken care of by a coordinate transformation in this analysis, for the
second one the lens configuration is assumed to be constant in time.

For the purpose of this analysis, it was desirable to mimic the
analytical behaviour of a simple fold as much as possible, for
comparing the numerical result with the analytical one, therefore, the
path of the source was chosen, so that it intersects the border of the
caustic perpendicularly, and on a point where the border is a fold
caustic.  We choose the line AC in Figure~\ref{fig:magnification_map}
and consider its two caustic-crossing segments separately:
\begin{itemize}
\item ``path~I'' runs from C to B;
\item ``path II'' runs from A to B.
\end{itemize}
The magnification along paths I and II are very similar, but not
identical.

The mock lightcurve used a crescent source (as explained in
\S\ref{subsec:crescent}) with parameters
\begin{equation}
   \left(R_p, \frac{R_n}{R_p}, \frac{a}{R_p}, \frac{b}{R_p}\right) =
   (1.31, 0.4, 0, 0.3)
\label{eqn:cp}
\end{equation}
corresponding to $r_{1/2}=1$.  Unit length has 30 pixels on the
magnification map.  The source was carried for 250 pixels along part
of path~I, giving 251 brightness values.  Gaussian noise at the level
of 4\% of the current brightness was then added to give the mock
lightcurve.  If we assume the brightness measurements are nightly, and
the transverse speed is 600~km/s, $r_{1/2}$ comes to 24~light-minutes,
which is plausible for a quasar.

\subsection{Model-fitting results}

We now proceed to fit the mock lightcurve using MCMC, to recover the
source parameters.

Note that fitting involves three nuisance parameters, namely the
brightness normalisation and the start- and endpoints on the path.
The latter two are equivalent to considering the location of the
caustic with respect to the starting point and the speed of the
source.  The fitting method automatically marginalises over all the
nuisance parameters.

Figure \ref{fig:cc_forward} shows a fit to the mock lightcurve and
recovery of $r_{1/2}$ and $R_n/R_p$.

Figure \ref{fig:cc_backward} shows the result when a different caustic
is used to fit, than was used to generate the mock lightcurve.  This
mimics the unavoidable systematic error of not knowing the caustic
exactly.  The fit is slightly degraded, but still acceptable.

Figure \ref{fig:gc_forward} shows the result of attempting to a
Gaussian source.  There is a small but statistically very significant
misfit because a Gaussian curve cannot reproduce the part of the
lightcurve with the slight flattening and then a second sub-peak
(cf.~Figure~\ref{fig:char_points}) as the dark part of the crescent
crosses the caustic.

\section{Discussion}\label{sec:discussion}

In the current paper, we simulate and study the resulting microlensing lightcurves of geometric crescent-shaped sources 
and compare them with the microlensing lightcurves of other simple mathematically describable source profiles. 
In order to mimic the behaviour of the flux of light from the source in the proximity of a fold caustic, we make use of the simple approximation described in equation (5). 
The equation would exhaustively describe the magnification map and offer a good universal approximation for the 
particular microlensing regime that we consider. 
Namely, the shape of the caustic boundary in the proximity of the source in the respective plane can be approximated by
 a line due to reason that the local radius of curvature of caustic is orders of magnitude greater than the half-light
 radius of the studied source. 
%In particular cases in which the previously mentioned approximation loses its validity, the impact on the quality of the 
%lightcurves is not evenly distributed. The shape of the lightcurve will be maintained. 
%The data points corresponding to the source position before and during the overlapping of the caustic will be affected by
% smaller errors than the data points corresponding to later times. \\

The first two source profiles that we consider are the uniform disk and symmetric Gaussian source. 
Both of them can be described by a half-light radius ($r_{1/2}$) and a total unlensed light flux ($S_0$). 
With the two parameters constrained the one-dimensional profiles, as well as the lightcurves of the 
two source, are completely determined, since no free parameter remains.

The previous statement does not hold for a crescent source.
In the case of the crescent source there are in total five parameters: the integrated flux of the source $S_0$,
 the radii of the bright and dark disks and the displacement of the centers of the two disks on the axes perpendicular and parallel to the caustic.

%Two of the parameters can be reduced by expressing the results in 
%terms of $S_0$ and $r_{1/2}$. 
%The later being determinable for any set of parameters 
%$R_p, R_n$ and $a^2+b^2$. Moreover, one of the displacement 
%parameters $b$ has no impact on the one-dimensional profile 
%of the source that results from the projection of the source
% image on an axis perpendicular to the caustic.

Since the one-dimensional source profile that corresponds to an 
axis perpendicular to the caustic contains exhaustively all the 
information regarding the source that can be revealed by the lightcurve, 
the value of the crescent's disks centers displacement parallel to the caustic 
does not have an effect on the shape of the lightcurve.

Nevertheless, the respective displacement is relevant for the calculation of
of the half-light radius.  It's qualitative effect is to 
decrease the value of the
half-light radius when the absolute value of the parameter is
increased.  In contrast to the simpler source shapes discussed before the
crescent has two additional relevant parameters: the radius of the dark 
disk ($R_n$) and the center displacement in the direction
 perpendicular to the caustic $a$.
Figure 4 reveals that the lightcurve of a crescent source has more
visible features than the other two light-curves corresponding to the
disc and Gaussian shape. The parameters $R_n$ and $a$ have strong
influences on the shape of the microlensed lightcurve as can be seen
in Figures~\ref{fig:a_var}, \ref{fig:lightcurve_crescent} and
\ref{fig:lightcurve_crescent_back}.  Moreover, the one-dimensional
source profile corresponding to the direction perpendicular to the
caustic reveal four characteristic points. The overlap of each of
these points with the caustic leaves visible features on the
lightcurve at the corresponding instances of time. In timely order,
the instances correspond to the start of the overlap between the
caustic and the bright disk, the start of the overlap between the
caustic and the dark disk, the end of the overlap between the caustic
and dark disk and finally the end of the overlap between the caustic
and the bright disk.

With the different source profiles and their corresponding lightcurves 
studied we can change our point of view of the system to that of an observer. 
The observer would basically detect only the lightcurve of such a source.
 As described in section 4.3 the timing of the onset and 
offset of the previously 
described periods can be used to estimate the values of the radii 
and one of the displacement parameters when assuming a geometric crescent. 
All quantities can be estimated in terms of the relative velocity 
of the source in a direction perpendicular to the caustic.

Furthermore, a simulated image of M87 presented in
 \citep{2012MNRAS.421.1517D} has been microlensed (figure 9). 
On the resulting lightcurve the instances 
corresponding to the start and end of the black hole 
shadow and caustic overlap were distinguishable.

In the case of a high-quality lightcurve with 
insignificant noise and measurement errors, 
the parameters can be obtained 
by simply identifying the characteristic 
instances of time without making use of 
the actual values of the magnification
 map. If the effect of the errors and the 
noise distorts the magnification time function enough so that the 
characteristic epochs are not identifiable with the characteristic 
periods still visible, the boundaries of the periods 
can be roughly estimated. Furthermore, if direct estimates of 
the parameters cannot be obtained we propose the use of a strong
statistical tool such as Markov-Chain Monte Carlo.
 In Section 6  we have studied the possibility of identifying a crescent 
source and the possibility of recovering the 
respective parameters. As magnification map, we have used a complex numerical 
one generated with the microlensing code by \cite{1999A&A...346L...5W}. 
In addition, we have added to the signal a Gaussian 
noise with an SNR of 1.6.  In the experiment,
 we have considered all nine combinations of
 original source profiles and assumed fitting source models.
Effectively we have fitted using MCMC all three sources with all three assumed fitting models. The results of the experiment allowed us to build a 
procedure for distinguishing the shape of the source assuming that one of the three models we have considered is a good approximation.  
The procedure would be useful for observers that endeavour to gain more information about an unresolved source for which they can study 
the microlensing lightcurve.  As a first step in the procedure, one should first attempt to use MCMC with a Gaussian model assumption. 
If either the value of the $\chi^2$ or the number of rejected datapoints is large then the next step is to change the assumed source model to 
a uniform disc and redo the fitting. Finally, if the uniform disc assumption is rejected as well then the fitting should be done with a 
crescent source model assumption. If at each of the three steps the data rejects the model then the source cannot be approximated by any of the three models. 
Otherwise, if at one of the steps the data does not reject the model then the respective model is a good approximation.

The previously mentioned abstract parameters can be related to
physical quantities specific to the central region of a quasar.  As
such the luminous region would correspond to the bright accretion disc
that surrounds the black hole. The later's gravity would cause a
shadow in the bright region limited by the extent of the event horizon
of the black hole. Therefore, the radius of the bright disk would
provide an estimate of the size of the accretion disk and the radius
of the dark disc would provide and estimate of the gravitationally
magnified Schwarzschild radius of the black hole $R_{S}^{magnified} =
\Delta t_{dark} \cdot v_p$.  By $R_{S}^{magnified}$ we refer to the
apparent Schwarzschild radius which is larger than the real value at
large distances due to the black hole's own gravity.  Moreover, the
gravitationally magnified value of the Schwarzschild radius is a
monotonic function of the black hole's mass. Therefore, it can be used
to estimate the mass of the black hole if it was not rotating. In the
previous expression, the $\Delta t_{dark}$ denotes the period of
overlap between the black hole shadow and fold caustic.  $v_p$ denotes
the component of the relative velocity of the source and fold which is
perpendicular to the caustic.  The respective velocity is an unknown,
though it can be constrained on a case by case basis to an order of
magnitude or even better. This would require the study of the dynamics
of the stellar structure which contains the gravitational lens.  A
better estimate of the relative velocity would facilitate a better
estimate of the effective non-rotating black hole mass associated with
the black hole shadow.

The parameters whose values cannot be determined due to the 
loss of information from the directions parallel to the 
caustic could be obtained in the eventuality in which 
the same source crosses multiple caustics that are not parallel. 
Multiple crossing of caustics can reveal details of 
the one-dimensional flux profile corresponding to multiple distinct 
directions which would allow the reconstruction of the 
two-dimensional profile analogous to the process through 
which an image of a CT scan is obtained.

\textit{Author contributions}\\
Prasenjit Saha provided the original idea and plan for the research project as well as multiple contributions to the analysis 
and manuscript preparation. Mihai Tomozeiu simulated and studied the ideal behaviour of the microlensing lightcurves for the 
different source profiles discussed and prepared the manuscript.
Joachim Wambsganss contributed to the research planning and provided the numerical code used by Manuel Rabold to create 
the magnification map and corresponding lightcurves used in the MCMC analysis performed by Irshad Mohammed in the last 
part of the presented work. Both Manuel Rabold and Irshad Mohammed had large contributions in writing the "Fitting Mock Data" section.

\section{Acknowledgement}
IM is supported by Fermi Research Alliance, LLC under Contract No. De-AC02-07CH11359 with the United States Department of Energy.
J.W. would like to acknowledge and thank the Pauli Center for Theoretical Studies of ETH Zurich and University of Zurich for 
generous support during the Schrödinger visiting professorship in 2013.

\bibliographystyle{mn2e}

\def\apj{ApJ}
\def\apjl{ApJL}
\def\aj{AJ}
\def\mnras{MNRAS}
\def\aap{A\&A}
\def\nat{nature}
\def\araa{ARAA}
\def\pasa{PASA}
\bibliography{heap}

\end{document}